---
title: "Bellabeat Project"
author: "Stephanie"
date: "2024-02-27"
output: html_document
---
## Load Packages
```{r}
library(tidyverse)
library(ggplot2)
library(conflicted)
library(skimr)
library(janitor)
library(lubridate)
```

## Load and Preview datasets
```{r}
head(dailyActivity_merged)
head(hourlySteps_merged)
head(sleepDay_merged)
```


## Verify structure of datasets
```{r}
str(dailyActivity_merged)
str(hourlySteps_merged)
str(sleepDay_merged)
```

## Verify number of variables
```{r}
n_distinct(dailyActivity_merged$Id)
n_distinct(hourlySteps_merged$Id)
n_distinct(sleepDay_merged$Id)
```

# CLEAN DATA
## Cheak For Duplicates
```{r}
sum(is.na(dailyActivity_merged))
sum(is.na(hourlySteps_merged))
sum(is.na(sleepDay_merged))
```

## Remove duplicates from sleepDay_merged
```{r}
sleepDay_merged <- sleepDay_merged[!duplicated(sleepDay_merged), ]
nrow(sleepDay_merged[duplicated(sleepDay_merged), ])
```

## Change column names to lowecases
```{r}
clean_names(dailyActivity_merged)
dailyActivity_merged <- rename_with(dailyActivity_merged, tolower)
clean_names(hourlySteps_merged)
hourlySteps_merged <- rename_with(hourlySteps_merged, tolower)
clean_names(sleepDay_merged)
sleepDay_merged <- rename_with(sleepDay_merged, tolower)
```

## Convert date string of hourlySteps_merged to date time and add a column for hours
```{r}
hourlySteps_merged$activityhour=as.POSIXct(hourlySteps_merged$activityhour,format="%m/%d/%Y %I:%M:%S %p")
hourlySteps_merged$hour <-  format(hourlySteps_merged$activityhour,format= "%H")
head(hourlySteps_merged)
```

## Add week column to dailyActivity_merged
```{r}
dailyActivity_merged <- dailyActivity_merged %>% 
  mutate( weekday = weekdays(as.Date(ActivityDate, "%m/%d/%Y")))
```

## Merge dailyActivity_merged and sleepDay_merged
```{r}
daily_act_sleep <- merge(dailyActivity_merged, sleepDay_merged, by = "id")
```

## Verify data in merged datasets
```{r}
sum(is.na(daily_act_sleep))
sum(duplicated(daily_act_sleep))
n_distinct(daily_act_sleep$Id)
colnames(daily_act_sleep)
```
## Order days of the week from sunday to satarday
```{r}
daily_act_sleep$weekday <- ordered(daily_act_sleep$weekday, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
daily_act_sleep[order(daily_act_sleep$weekday), ]
```



# ANALYZE
## For the daily_act_sleep dataframe summary
```{r}
daily_act_sleep %>%
  select(totalsteps,
         totaldistance,
         veryactiveminutes,
         sedentaryminutes,
         calories,
         totalsleeprecords,
         totalminutesasleep,
         totaltimeinbed) %>%
  summary()
```

## Summary of hourlySteps_merged
```{r}
hourlySteps_merged %>%
  select(steptotal) %>%
  summary()
```

## Let visualize steps per week
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=totalsteps, fill=weekday))+ 
  geom_bar(stat="identity", fill="pink")+
  labs(title="Steps per weekday", y="Total Steps")
```

## Let visualize calories burnt during the week
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=calories, fill=weekday))+ 
  geom_bar(stat="identity", fill="blue")
```

## Let visualize active minutes
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=veryactiveminutes, fill=weekday))+ 
  geom_bar(stat="identity", fill="Orange")+
  labs(title="Active Minutes", y="Active Minutes")
```


## Let visualize sedentary minutes
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=sedentaryminutes, fill=weekday))+ 
  geom_bar(stat="identity", fill="red")+
  labs(title="Sedentary Minutes", y="Sedentary Minutes")
```

## Let visualize total minutes asleep
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=totalminutesasleep, fill=weekday))+ 
  geom_bar(stat="identity", fill="purple")+
  labs(title="Total Minutes Asleep During the Week", y="Total Minutes Asleep")

```

## Let visualize total distance
```{r}
ggplot(data=daily_act_sleep, aes(x=weekday, y=totaldistance, fill=weekday))+ 
  geom_bar(stat="identity", fill= "green")+
  labs(title = "Total distance covered", y= "Total Distance")
```

## Let visualize hourly steps
```{r}
ggplot(data=hourlySteps_merged, aes(x=hour, y=steptotal, fill=hour))+
  geom_bar(stat="identity")+
  labs(title="Hourly Steps")
```

# Visualize total steps and calories
```{r}
ggplot(data = daily_act_sleep, aes(x = totalsteps, y = calories)) + 
  geom_point() + geom_smooth() + labs(title ="Total Steps vs. Calories")
```










